---
title: "Class 11"
format: pdf
---

#Section 1. Proportion og G/G in a population



```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378-4.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale.

How many samples do we have?
>Q13. Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
```{r}
nrow(expr)
```
```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```


Let's make a boxplot

```{r}
ggplot(expr) + aes(geno, exp, fill=geno) + geom_boxplot(notch=TRUE)
```

>Q14.Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3? 

Looks like having a G/G in this location is associated with having a reduced expression of ORMDL3. According to the boxplot you could infer that A/A has a higher expression value than G/G. 










