---
title: "Class 10: Halloween Mini-project"
format: pdf
---

#1. Importing candy data

```{r}
candy_file <- "class 10 .txt"
candy = read.csv(candy_file, row.names =1)
head(candy)
```


```{r}
nrow(candy)
```
>Q1. How many different candy types are in this dataset?

85

```{r}
sum(candy [,2])
```
>Q2. How many fruity candy types are in the dataset?

38

#2. What is your favorate candy?

```{r}
candy["Twix", ]$winpercent
```
>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy is Twix and the winpercent value is 81.64291.

```{r}
candy["Kit Kat", ]$winpercent
```
>Q4. What is the winpercent value for “Kit Kat”?

76.7686

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

49.6535

```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The hist column and the skim_variable column are different because they are not numeric. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

0 does not have chocolate and 1 does have chocolate. 

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```
>Q9. Is the distribution of winpercent values symmetrical?

The distribution of winpercent values is not symmetrical the histogram is skewed to the right. 

>Q10. Is the center of the distribution above or below 50%?

It is above 50%

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])
```

```{r}
mean(candy$winpercent[as.logical(candy$fruity)])
```
On average chocolate candy is higher ranked than fruity candy. 



>Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)],candy$winpercent[as.logical(candy$fruity)])
```
It is statistically significant because the p-value is below 0.05. 

```{r}
head(candy[order(candy$winpercent),], n=5)
```
```{r}
library(dplyr)
```
#3. Overall Candy Rankings

>Q13. What are the five least liked candy types in this set?

```{r}
candy %>% arrange(winpercent) %>% head(5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
candy %>% arrange(-winpercent) %>% head(5)
```
>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) + geom_col()
```


```{r}
my_cols <- ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) + geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
  
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

> Q17. What is the worst ranked chocolate candy?

Sixlets

>Q18. What is the best ranked fruity candy?

Starburst

#4. Taking a look at pricepercent


```{r}
ord <- order(candy$pricepercent, decreasing = FALSE)
head( candy[ord,c(11,12)], n=5 )
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Tootsie Roll Midgies yields the highest winpercent for the lowest price percent meaning it offers the most bang for your buck. 

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

Nik L Nip, Nestle Smarties, Ring pop, Hershey's Krackel, and Hershey's Milk Chocolate are the most expensive. Nik L Nip is the least popular.

#5 Exploring the correlation structure

```{r}
library(corrplot)
```
```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Chocolate and fruity variables are the most anti-correlated.

>Q23. Similarly, what two variables are most positively correlated?

Chocolate and Bar are most positively correlated. 

#6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1:2])
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 1000)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```


```{r}
library(plotly)
```

```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus are picked up strongly by PC1 in the positive direction. "Picked up strongly in the positive direction" means that there is a strong and positive association between the variables, such that an increase in one variable is associated with an increase in the other variable.This makes sense because fruity candy is usually hard and tends to come pluribus.



